### 2022-08-13 draft.tex
Deploying to gh-pages from @ FrancisBanville/ms_maxent_networks@fe5f2b2b1a820012f671a6765f220776b569f59a 🚀

A variety of measures of the structure of ecological networks have been
used to describe the organization of species interactions in a
biological community (Delmas et al. 2019). These measures provide
valuable information on the functioning of ecosystems and their
responses to environmental change (e.g., Pascual and Dunne 2006; Gómez,
Perfectti, and Jordano 2011). For instance, Bascompte et al. (2003)
suggest that plant--pollinator and seed-disperser networks have a highly
nested structure that can promote species persistence. Another example,
in food webs, shows that a high connectance can promote the robustness
of the system to species lost (Dunne, Williams, and Martinez 2002).
However, despite the growing literature on the ecological implications
of network structure, the association between many of these measures
impedes our ability to fully understand what drives the structure and
behavior of ecological networks. In particular, nestedness and
modularity are strongly associated in ecological networks (Fortuna et
al. 2010), and network connectance has been shown to be an important
driver of many other emerging network properties (Timothée Poisot and
Gravel 2014). In light of these observations, it is difficult to assess
whether attributed effects of given properties are the artifacts of
other, perhaps simpler, measures.

One way to tackle this issue is first to recognize that food webs and
other ecological networks are constrained systems. In other words, the
space of possible network configurations shrinks as we know more about a
network structure. For example, there is a finite number of networks
with specified numbers of nodes and edges. Indeed, the structure of
ecological networks is first and above all constrained by the number of
species, or nodes, present. Food webs with high species richness
typically have a lower connectance (MacDonald, Banville, and Poisot
2020) than smaller networks. This is because the number of realized
interactions in empirical food webs scales slower than the number of
possible species pairs (MacDonald, Banville, and Poisot 2020). As shown
by Timothée Poisot and Gravel (2014), connectance itself can constrain
different aspects of network structure such as the degree distribution
(i.e.~the probability distribution of the number of interspecific
interactions realized by a species). Other measures, such as the maximum
trophic level, can also constrain the space of feasible networks.

Prior knowledge on the structure of ecological networks is thus
especially useful in the current context of data scarcity about species
interactions (Jordano 2016). Indeed, this Eltonian shortfall (Hortal et
al. 2015) can be partially alleviated using known information about an
ecological network. As suggested by Strydom et al. (2021), network
structure can be used to improve the prediction of pairwise species
interactions when data is lacking by constraining the space of feasible
networks. Similarly, partial knowledge on the structure of an ecological
network can also be used to predict others of its properties by
constraining their range of possible values. This is important given
that many aspects of network structure cannot be measured empirically
without data on pairwise species interactions, a prevailing situation
worldwide (Timothée Poisot et al. 2021).

@@

We tested our analytical and heuristic MaxEnt models against open
food-web data queried from three different sources and integrated into
what we call our \emph{complete dataset}. These sources include (1)
terrestrial and aquatic food webs sampled globally and archived on the
ecological interactions database Mangal; (2) freshwater stream food webs
from New Zealand; and (3) aquatic food webs from Tuesday Lake (Michigan,
United States). First, all food webs archived on \texttt{mangal.io}
(Timothee Poisot et al. 2016; Banville, Vissault, and Poisot 2021) were
directly queried from the database (\(N = 235\)). Most ecological
networks archived on Mangal are multilayer networks, i.e.~networks that
describe different types of interactions. We kept all networks whose
interactions were mainly of predation and herbivory types, and removed
the largest network (\(S = 714\)) for computational efficiency reasons.
Then, to this set we added food webs from two different sources: the New
Zealand dataset (\(N = 21\); J. P. Pomeranz et al. 2018) and the Tuesday
Lake dataset (\(N = 2\); Cohen, Jonsson, and Carpenter 2003). Our
complete dataset thus contained a total of \(257\) food webs. Of the New
Zealand and Tuesday Lake datasets, \(19\) networks had data on species'
relative abundances that were used in the neutral model (Box 2). These
\(19\) food webs constitute our \emph{abundance dataset}, which is a
subset of our complete dataset.

@@

We compared the performance of both heuristic MaxEnt models in
predicting these measures to the one of the null
(tbl.~\ref{tbl:measures_all}) and neutral models
(tbl.~\ref{tbl:measures_abund}). Overall, we found that the models based
on the joint degree sequence (i.e.~the type II null and heuristic MaxEnt
models) reproduced the structure of empirical food webs much better than
the ones based on connectance (i.e.~the type I null and heuristic MaxEnt
models). This suggests that the predictive power of connectance might be
more limited than what was previously suggested (Timothée Poisot and
Gravel 2014). On the other hand, the neutral model of relative
abundances was surprisingly good at predicting the maximum trophic level
and the network diameter. However, with the exception of the network
diameter, the type II heuristic MaxEnt model was better at predicting
network structure than the neutral model for most measures considered.
This might be because, although neutral processes are important, they
act in concert with niche processes in determining species interactions
Canard et al. (2014). The joint degree sequence encodes information on
both neutral and niche processes because the number of prey and
predators a species has is determined by its relative abundance and
biological traits. These results thus show that having information on
the number of prey and predators for each species substantially improves
the prediction of food-web structure, both compared to models solely
based on connectance and to the ones solely based on species relative
abundances.

@@

\leavevmode\hypertarget{ref-Poisot2016ManMakb}{}%
Poisot, Timothee, Benjamin Baiser, Jennifer A. Dunne, Sonia Kefi,
Francois Massol, Nicolas Mouquet, Tamara N. Romanuk, Daniel B. Stouffer,
Spencer A. Wood, and Dominique Gravel. 2016. {``Mangal - Making
Ecological Network Analysis Simple.''} \emph{Ecography} 39 (4): 384--90.
\url{https://doi.org/10.1111/ecog.00976}.

>>>>

A variety of measures of the structure of ecological networks have been
used to describe the organization of species interactions in a
biological community (Delmas et al. 2019). These measures provide
valuable information on the functioning of ecosystems and their
responses to environmental change (e.g., Pascual and Dunne 2006; Gómez,
Perfectti, and Jordano 2011). For instance, Bascompte et al. (2003)
suggest that plant--pollinator and seed-disperser networks have a highly
nested structure that can promote species persistence. Another example,
in food webs, shows that a high connectance can promote the robustness
of the system to species lost (Dunne, Williams, and Martinez 2002).
However, despite the growing literature on the ecological implications
of network structure, the association between many of these measures
impedes our ability to fully understand what drives the structure and
behavior of ecological networks. In particular, nestedness and
modularity are strongly associated in ecological networks (Fortuna et
al. 2010), and network connectance has been shown to be an important
driver of many other emerging network properties (Poisot and Gravel
2014). In light of these observations, it is difficult to assess whether
attributed effects of given properties are the artifacts of other,
perhaps simpler, measures.

One way to tackle this issue is first to recognize that food webs and
other ecological networks are constrained systems. In other words, the
space of possible network configurations shrinks as we know more about a
network structure. For example, there is a finite number of networks
with specified numbers of nodes and edges. Indeed, the structure of
ecological networks is first and above all constrained by the number of
species, or nodes, present. Food webs with high species richness
typically have a lower connectance (MacDonald, Banville, and Poisot
2020) than smaller networks. This is because the number of realized
interactions in empirical food webs scales slower than the number of
possible species pairs (MacDonald, Banville, and Poisot 2020). As shown
by Poisot and Gravel (2014), connectance itself can constrain different
aspects of network structure such as the degree distribution (i.e.~the
probability distribution of the number of interspecific interactions
realized by a species). Other measures, such as the maximum trophic
level, can also constrain the space of feasible networks.

Prior knowledge on the structure of ecological networks is thus
especially useful in the current context of data scarcity about species
interactions (Jordano 2016). Indeed, this Eltonian shortfall (Hortal et
al. 2015) can be partially alleviated using known information about an
ecological network. As suggested by Strydom et al. (2021), network
structure can be used to improve the prediction of pairwise species
interactions when data is lacking by constraining the space of feasible
networks. Similarly, partial knowledge on the structure of an ecological
network can also be used to predict others of its properties by
constraining their range of possible values. This is important given
that many aspects of network structure cannot be measured empirically
without data on pairwise species interactions, a prevailing situation
worldwide (Poisot et al. 2021).

@@

We tested our analytical and heuristic MaxEnt models against open
food-web data queried from three different sources and integrated into
what we call our \emph{complete dataset}. These sources include (1)
terrestrial and aquatic food webs sampled globally and archived on the
ecological interactions database Mangal; (2) freshwater stream food webs
from New Zealand; and (3) aquatic food webs from Tuesday Lake (Michigan,
United States). First, all food webs archived on \texttt{mangal.io}
(Poisot et al. 2016; Banville, Vissault, and Poisot 2021) were directly
queried from the database (\(N = 235\)). Most ecological networks
archived on Mangal are multilayer networks, i.e.~networks that describe
different types of interactions. We kept all networks whose interactions
were mainly of predation and herbivory types, and removed the largest
network (\(S = 714\)) for computational efficiency reasons. Then, to
this set we added food webs from two different sources: the New Zealand
dataset (\(N = 21\); J. P. Pomeranz et al. 2018) and the Tuesday Lake
dataset (\(N = 2\); Cohen, Jonsson, and Carpenter 2003). Our complete
dataset thus contained a total of \(257\) food webs. Of the New Zealand
and Tuesday Lake datasets, \(19\) networks had data on species' relative
abundances that were used in the neutral model (Box 2). These \(19\)
food webs constitute our \emph{abundance dataset}, which is a subset of
our complete dataset.

@@

We compared the performance of both heuristic MaxEnt models in
predicting these measures to the one of the null
(tbl.~\ref{tbl:measures_all}) and neutral models
(tbl.~\ref{tbl:measures_abund}). Overall, we found that the models based
on the joint degree sequence (i.e.~the type II null and heuristic MaxEnt
models) reproduced the structure of empirical food webs much better than
the ones based on connectance (i.e.~the type I null and heuristic MaxEnt
models). This suggests that the predictive power of connectance might be
more limited than what was previously suggested (Poisot and Gravel
2014). On the other hand, the neutral model of relative abundances was
surprisingly good at predicting the maximum trophic level and the
network diameter. However, with the exception of the network diameter,
the type II heuristic MaxEnt model was better at predicting network
structure than the neutral model for most measures considered. This
might be because, although neutral processes are important, they act in
concert with niche processes in determining species interactions Canard
et al. (2014). The joint degree sequence encodes information on both
neutral and niche processes because the number of prey and predators a
species has is determined by its relative abundance and biological
traits. These results thus show that having information on the number of
prey and predators for each species substantially improves the
prediction of food-web structure, both compared to models solely based
on connectance and to the ones solely based on species relative
abundances.

@@

\leavevmode\hypertarget{ref-Poisot2016ManMakb}{}%
Poisot, Timothée, Benjamin Baiser, Jennifer A. Dunne, Sonia Kefi,
Francois Massol, Nicolas Mouquet, Tamara N. Romanuk, Daniel B. Stouffer,
Spencer A. Wood, and Dominique Gravel. 2016. {``Mangal - Making
Ecological Network Analysis Simple.''} \emph{Ecography} 39 (4): 384--90.
\url{https://doi.org/10.1111/ecog.00976}.
